A meta-analysis of fungicide spray schedules to minimise yield loss from powdery mildew in Australian mungbean

### 2020-08-28 03_Preliminary_analysis.Rmd
Clean up preliminary analysis, more concise code

## Explore and Visualise Data
Each row provides the mean yield for each experimental treatment.

The data explored in the `PM_MB_dat` data frame is a list of mean values from each treatment in each respective experiment.
These preliminary analyses will describe the data and then investigate any change in mungbean yields which is attributable to the fungicide treatment. 
### Descriptive statistics
#### Spatio-temporal variation

@@

```{r FinalTrialsForInclusion}
tmp2 <- PM_MB_dat %>%
   filter(fungicide_ai == "tebuconazole" |
             fungicide_ai == "propiconazole")%>%
   distinct(trial_ref, location, year)%>%
   arrange(year)
   table(tmp1$location)
```

@@

What is the variation in yield over all the trials (including non-DMI fungicide trials) in the no spray control and fungicide treated plots.  
```{r ControlYield_AllTrials}
PM_MB_dat %>%
   filter(fungicide_ai == "control")%>%
   summarise(mean_yield = mean(grain_yield.t.ha.),
             minimum_yield = min(grain_yield.t.ha.),
             maximum_yield = max(grain_yield.t.ha.),
             yield_stdev = sd(grain_yield.t.ha.))

@@

PM_MB_dat %>%
   filter(trial_ref %in% DMITrials)%>%
   filter(fungicide_ai == "control" )%>%
   summarise(mean_yield = mean(grain_yield.t.ha.*1000),
             minimum_yield = min(grain_yield.t.ha.*1000),
             maximum_yield = max(grain_yield.t.ha.*1000),
             yield_sterr = sd(grain_yield.t.ha.*1000)/(sqrt(length(grain_yield.t.ha.))))

@@

Mean yields in DMI fungicide treated plots
```{r Yield_DMITrials}
PM_MB_dat %>%
   filter(fungicide_ai == "tebuconazole" |
             fungicide_ai == "propiconazole")%>%
   summarise(mean_yield = mean(grain_yield.t.ha.*1000),
             minimum_yield = min(grain_yield.t.ha.*1000),
             maximum_yield = max(grain_yield.t.ha.*1000),
             yield_sterr = sd(grain_yield.t.ha.*1000)/(sqrt(length(grain_yield.t.ha.))))

@@

```{r}
mpg %>%
   group_by(manufacturer,drv,cyl)%>%
   filter(n() >= 2)%>%
   summarise(mean_Sp = mean(cty),
             stdev_h = sd(cty),
             n_group = n())

data()

mpg

Later in this document we will also consider the impact of fungicide treatments on disease incidence/severity.
Mean powdery mildew severity was recorded at the end of each growing season.
Some experiments recorded disease severity throughout the season and these were summarised with AUDPC.
AUDPC for trials which only provided final powdery mildew severity were calculated using the record of 'first sign of disease' and final assessment.

@@

Let's investigate the difference in the effect of the type of fungicide on proportion of yield saved, but first we need to calculate yield saved relative to the no spray control.

Now we have calculated proportional yield savings we can inspect the difference between the fungicide active ingredients

@@

```{r RS_severity}
# Which row spacing leads to the higher disease severity
PM_MB_dat %>%
   filter(year == 2017 |
             year == 2018) %>%
   filter(fungicide_ai == "control") %>%
   ggplot(aes(y = PM_final_severity, x = factor(row_spacing))) +
   geom_boxplot(fill = usq_cols("usq charcoal"), alpha = 0.5) +
   ylab("Final severity rating") +
   xlab("Row spacing (m)") +
   ggtitle("Powdery mildew variation between different row spacing")
```

@@

```{r RS_yield}
# Which row spacing leads to the higher yield potential
PM_MB_dat %>%
   filter(year == 2017 |
             year == 2018) %>%
   ggplot(aes(
      y = as.numeric(grain_yield.t.ha.),
      x = as.factor(row_spacing),
      colour = location
   )) +
   geom_jitter(width = 0.1) +
   ggtitle("Grain yield results for different row spacing at three locations") +
   xlab("Row spacing (m)") +
   ylab("Grain yield (t/ha)") +
   scale_colour_usq()
```

@@

### AUDPC visualise raw data

First let's check if we have all the AUDPC observations by searching for trials with mission AUDPC.

```{r AUDPC_check}
PM_MB_dat %>%
   filter(is.na(AUDPC_m)) %>%
   distinct(trial_ref, location, year,AUDPC_m, first_sign_disease, final_assessment)%>%
   kable(align = "c")
```

The trials `r unique(PM_MB_dat[is.na(PM_MB_dat$AUDPC_m),"trial_ref"]) ` don't have an estimate of AUDPC and need to be filled in. 
The following script estimates AUDPC from the final disease severity rating and the first sign of disease information.

```{r SingleObsAUDPC}
source(here("R/AUDPC_for_single_obs.R"))
```

      _How well does 'area under the disease progress curve' (AUDPC) predict grain yield?_

   and

      _Does spray_management scenario mitigate yield loss through reducing AUDPC?_

We would expect to see grain yield decline with increasing AUDPC.
Let's just plot yield vs AUDPC.

```{r Yield_vs_AUDPC}

plot(grain_yield.t.ha. ~ AUDPC_m, data = PM_MB_dat)
abline(lm(grain_yield.t.ha. ~ AUDPC_m, data = PM_MB_dat))

There appears to be a weak positive relationship between the two continuous variables.
Does a linear model confirm this?

```{r AUDPClm}
summary(lm(grain_yield.t.ha. ~ AUDPC_m, data = PM_MB_dat))
```

This is not what we would expect and the low R squared value of `r round(summary(lm(grain_yield.t.ha. ~ AUDPC_m, data = PM_MB_dat))$adj.r.squared,3)` perhaps there are additional factors influencing this relationship.

Let's tease out variation due to trial location, season, row spacing, and other seasonal variables are preventing us from seeing the expected relationship.
Let's do the plot again for each trial where there are no missing values.
Note: The plots are ordered by mean yield of the trial.

```{r AUDPCxTrial, fig.height= 10, fig.width=10}
mTrialYield <- PM_MB_dat %>%
   group_by(trial_ref) %>%
   mutate(TY = mean(grain_yield.t.ha., na.rm = TRUE)) %>%
   pull(TY)

PM_MB_dat$trial_ref <- reorder(PM_MB_dat$trial_ref, mTrialYield)

PM_MB_dat %>%
   filter(!is.na(grain_yield.t.ha.)) %>%
   mutate(Treatment = case_when(
      fungicide_ai == "control" ~ "control",
      TRUE ~ "FungicideTreated"))%>%
   ggplot(aes(y = grain_yield.t.ha., x = AUDPC_m))+
   geom_point(aes(color = Treatment))+
   facet_wrap(vars(trial_ref))+
   scale_y_continuous(limits = c(0,3))+
   geom_smooth(method = "lm", se = FALSE, linetype = 1, colour = "grey", size =1)+
   scale_colour_usq()

@@

glimpse(beans)

>>>>

## Explore and Visualise Data

The data explored in the `PM_MB_dat` data frame is a list of mean values from each treatment in each respective experiment.
These preliminary analyses will describe the data and then investigate any change in mungbean yields which is attributable to the fungicide treatment.

### Descriptive statistics  
#### Spatio-temporal variation

@@

```{r FinalTrialsForInclusion}
   table(tmp1$location)
```

@@

What is the variation in yield over all the trials (including non-DMI fungicide trials) in the no spray control and fungicide treated plots.

```{r ControlYield_AllTrials}
PM_MB_dat %>%
   filter(fungicide_ai == "control")%>%
   summarise(mean_yield = mean(grain_yield.t.ha.),
             minimum_yield = min(grain_yield.t.ha.),
             maximum_yield = max(grain_yield.t.ha.),
             yield_stdev = sd(grain_yield.t.ha.)) %>%
   kable(align = 'c',
   caption = "Average yield statistics for no spray contol plots"
)

@@

PM_MB_dat %>%
   filter(trial_ref %in% DMITrials)%>%
   filter(fungicide_ai == "control" )%>%
   summarise(mean_yield = mean(grain_yield.t.ha.*1000),
             minimum_yield = min(grain_yield.t.ha.*1000),
             maximum_yield = max(grain_yield.t.ha.*1000),
             yield_sterr = sd(grain_yield.t.ha.*1000)/(sqrt(length(grain_yield.t.ha.))))%>%
   kable(align = 'c',
   caption = "Average yield statistics for control plots in trials which used a DMI fungicide")

@@

Mean yields in DMI fungicide treated plots
```{r Yield_DMITrials}
PM_MB_dat %>%
   filter(fungicide_ai == "tebuconazole" |
             fungicide_ai == "propiconazole")%>%
   summarise(mean_yield = mean(grain_yield.t.ha.*1000),
             minimum_yield = min(grain_yield.t.ha.*1000),
             maximum_yield = max(grain_yield.t.ha.*1000),
             yield_sterr = sd(grain_yield.t.ha.*1000)/(sqrt(length(grain_yield.t.ha.))))%>%
   kable(align = 'c',
   caption = "Average yield statistics for DMI fungicide treated plots")

@@

@@

@@

```{r RS_severity}
# Which row spacing leads to the higher disease severity
PM_MB_dat %>%
   filter(year == 2017 |
             year == 2018) %>%
   filter(fungicide_ai == "control") %>%
   ggplot(aes(y = PM_final_severity, x = factor(row_spacing))) +
   geom_boxplot(fill = usq_cols("usq charcoal"), alpha = 0.5) +
   ylab("Final severity rating") +
   xlab("Row spacing (m)") +
   ggtitle("Powdery mildew severity between different row spacing across all trials")
```

@@

```{r RS_yield}
# Which row spacing leads to the higher yield potential
PM_MB_dat %>%
   filter(year == 2017 |
             year == 2018) %>%
   ggplot(aes(
      y = as.numeric(grain_yield.t.ha.),
      x = as.factor(row_spacing),
      colour = location
   )) +
   geom_jitter(width = 0.05) +
   ggtitle("Grain yield results for different row spacing at three locations") +
   xlab("Row spacing (m)") +
   ylab("Grain yield (t/ha)") +
   scale_colour_usq()
```

@@

@@

str(beans)
